---
title: "2024-11-18 course_part1"
author: "Eve Fiskebeck"
date: "`r Sys.Date()`"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Course part 1 - Laboratory

We selected 20 isolates :

-   6 isolates are ***Klebsiella spp.*** isolates
-   6 isolates are *Escherichia coli* isolates
-   8 isolates are *Salmonella spp.* isolates

```{=html}
<!-- 
TODO: I am a comment 

damned lab work it failed too much today 

-->
```
## Installation of packages

```{r}
#install.packages("here")
#install.packages("DiagrammeR")
```

ctrl + alt + I (make a cell - shortcut on the keyboard)

```{r}
angry_or_not <- "not"
```

gives content of the object (container)

```{r}
angry_or_not
```

Madelaine, are you angry or not? `r angry_or_not`. Something else

You can run inline text

# Working with your data


```{r}
here::here()
```

```{r}
path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")

path_data
```

```{r}
human_data <- 
  readr::read_csv2(path_data)
```
For `;`delimited files its read_csv2 function
For `,`delimited files its read_csv function

---

# Day 3

## Packages into memory 
```{r}
library(here)
library(tidyverse)
```


## Getting help - and what functions do 

F1 


```{r}
?library()
help("library")
```

## we get out data 
```{r}
path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")

path_data
```



```{r}
human_data <- 
  readr::read_csv2(path_data)
```





## Exploration of a data frame 

- Names of columns 
```{r}
colnames(human_data)
```

- getting the columns names into a vector for pasting and manipulations
- I can use a function within a function ! 
```{r}
dput(
  colnames(human_data)
  )
```


- explanation: c() is to make vectors

```{r}
test <- c("one", "2", 3)
test
test[3]
typeof(test[3])
typeof(test)
class(test)
```
All the elements of a vector are of the same `type` (text, double = Real, integeter, date ...)

```{r}
test2 <-c(1,2,3)
typeof(test2)
class(test2)
test2*test2
```
```{r}
test2 * c(1,2)
```


- deactivate the output in the markdown document: in the document settings
near knit button : only select `chunk output in the console`


- Overview of data and data types in your data

```{r}
glimpse(human_data)
str(human_data)
```

- viewing your data or better an abstract of your data. Notion of pipe 
```{r}

```

