---
title: "2024-11-18 course_part1"
author: "Eve Fiskebeck"
date: "`r Sys.Date()`"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Course part 1 - Laboratory

We selected 20 isolates :

-   6 isolates are ***Klebsiella spp.*** isolates
-   6 isolates are *Escherichia coli* isolates
-   8 isolates are *Salmonella spp.* isolates

```{=html}
<!-- 
TODO: I am a comment 

damned lab work it failed too much today 

-->
```
## Installation of packages

```{r}
#install.packages("here")
#install.packages("DiagrammeR")
```

ctrl + alt + I (make a cell - shortcut on the keyboard)

```{r}
angry_or_not <- "not"
```

gives content of the object (container)

```{r}
angry_or_not
```

Madelaine, are you angry or not? `r angry_or_not`. Something else

You can run inline text

# Working with your data


```{r}
here::here()
```

```{r}
path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")

path_data
```

```{r}
human_data <- 
  readr::read_csv2(path_data)
```
For `;`delimited files its read_csv2 function
For `,`delimited files its read_csv function

---

# Day 3

## Packages into memory 
```{r}
library(here)
library(tidyverse)
```

<!--
## Getting help - and what functions do 

F1 


```{r}
?library()
help("library")
```
-->


## we get out data 
```{r}
path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")

path_data
```



```{r}
human_data <- 
  readr::read_csv2(path_data)
```





## Exploration of a data frame 

- Names of columns 
```{r}
colnames(human_data)
```

- getting the columns names into a vector for pasting and manipulations
- I can use a function within a function ! 
```{r}
dput(
  colnames(human_data)
  )
```


- explanation: c() is to make vectors

```{r}
test <- c("one", "2", 3)
test
test[3]
typeof(test[3])
typeof(test)
class(test)
```
All the elements of a vector are of the same `type` (text, double = Real, integeter, date ...)

```{r}
test2 <-c(1,2,3)
typeof(test2)
class(test2)
test2*test2
```
```{r}
test2 * c(1,2)
```


- deactivate the output in the markdown document: in the document settings
near knit button : only select `chunk output in the console`


- Overview of data and data types in your data

```{r}
glimpse(human_data)
str(human_data)
```


- renaming columns names (easier to use)

```{r}
original_names <- colnames(human_data)
original_names
```

```{r}
library(tidyverse)
new_names <- str_replace_all(original_names, " ", "_" )
new_names
new_names[88]
```

- regex pattern 
```{r}
new_names2 <- str_replace_all(
  new_names, "[?/,;.*()-]", "_" ) %>%
  str_remove_all("(^_*)|(_*$)")



new_names2
new_names2[88]

str_replace_all(
  "emanuel /; ", 
  "[?/,;.*()-]", 
  "_" ) %>%
  str_replace_all(., " ", "_")
```


```{r}
colnames(human_data) <- new_names2
colnames(human_data)[88]
```


- selecting columns

```{r}
human_data %>% 
  select(INIKA_OH_TZ_ID, Age__yrs_, Gender) %>%
  head()
```


```{r}
human_data_selection <- 
  human_data %>%
  select(INIKA_OH_TZ_ID, Age__yrs, Gender, Enter_a_date, Region, District, 
         Specify_if_other_district, Sample, Season, Origin_of_sample, 
         Which_class_grade_are_you, 
         Who_is_your_caretaker, 
         If_others__mention, 
         What_is_your_occupation_and_or_of_your_caretaker, 
         Have_you_ever_heard_about_AMR, If_yes__how_did_you_get_this_information, 
         Have_you_or_your_children_used_any_antibiotics_at_any_time, 
         If_yes__where_did_you_get_these_drugs_from,
         If_it_was_drug_sellers_or_pharmacy__did_you_have_a_prescription_from_the_doctor_prescriber,
         GPS_coordinates_latitude, GPS_coordinates_longitude) 
```

# CLEAN 


```{r}
human_data <- 
  readr::read_csv2(path_data)
```


```{r}
new_columns <- 
  colnames(human_data) %>%
  str_replace_all(" ", "_") %>%
  str_replace_all( "[?/,;.*()-]", "_") %>%
  str_remove_all("(^_*)|(_*$)")
```

```{r}
colnames(human_data) <- new_columns
```

```{r}
human_data_selection <- 
  human_data %>%
  select(INIKA_OH_TZ_ID, Age__yrs, Gender, Enter_a_date, Region, District, 
         Specify_if_other_district, Sample, Season, Origin_of_sample, 
         Which_class_grade_are_you, 
         Who_is_your_caretaker, 
         If_others__mention, 
         What_is_your_occupation_and_or_of_your_caretaker, 
         Have_you_ever_heard_about_AMR, If_yes__how_did_you_get_this_information, 
         Have_you_or_your_children_used_any_antibiotics_at_any_time, 
         If_yes__where_did_you_get_these_drugs_from,
         If_it_was_drug_sellers_or_pharmacy__did_you_have_a_prescription_from_the_doctor_prescriber,
         GPS_coordinates_latitude, GPS_coordinates_longitude) 

human_data_selection %>% View()
human_data_selection %>% head()
```


->  viewing your data or better an abstract of your data. Notion of pipe and selecting columns 


```{r}
glimpse(human_data_selection)
str(human_data_selection)
```

# HERE 
- something does not make sense - empty column 

```{r}
human_data_selection <- 
  human_data_selection %>%
  select(-Specify_if_other_district) %>%
  mutate_at(vars(Age__yrs), as.integer) %>%
  mutate_if(is.character, factor) %>%
  mutate_at(vars(INIKA_OH_TZ_ID) , as.character) 
```

```{r}
str(human_data_selection)
```


```{r}
levels(test$Region)
```

- summary 
```{r}
summary(
  human_data_selection
  )
```


- looking at the content of single columns

- different values contained in columns 

- factors vs character 


- changing the types of the columns so they can be processed correctly 




```{r}
human_data_selection %>%
  filter(INIKA_OH_TZ_ID == "238")
```


```{r}
human_data_selection %>%
  filter(INIKA_OH_TZ_ID != "238") %>%
  filter(is.na(Gender))
  #summary() 

#%>%
# View()
```

